---
title: "Getting and cleaning data - Course Project"
author: "Remco Niggebrugge"
date: "01/14/2015"
output: html_document
---
 
 
**This document is for me to keep track**, I do not think a RMarkdown file should be posted as part of this assignment.

These are the steps I am going to take.

**0** Libraries

```{r libraries}
library(dplyr)
```

**1** Set working directory to that with all raw data files:

```{r setwd}
setwd("/home/remco/Desktop/data/small/")
```

**2** Read training files:

```{r read-train-set}
subject      <- read.table("./train/subject_train.txt")
activity     <- read.table("./train/y_train.txt")
measurements <- read.table("./train/X_train.txt", sep="")
features     <- read.table("./features.txt")
```

**3** Assign column names:

```{r assign-column-names}
names(subject)      <- "subject"
names(activity)     <- "activity"
names(measurements) <- features[,2]    ## use 2nd column of features, first column has index
```

**4** Now simply bind the columns (pasting data.frames left-to-right)

```{r binding-training}
train_frame <- cbind(subject, activity, measurements)
```

**5** Now repeat same steps for the test dataset:

```{r read-test-set}
subject      <- read.table("./test/subject_test.txt")
activity     <- read.table("./test/y_test.txt")
measurements <- read.table("./test/X_test.txt", sep="")
names(subject)      <- "subject"
names(activity)     <- "activity"
names(measurements) <- features[,2]    ## use 2nd column of features, first column has index

test_frame <- cbind(subject, activity, measurements)
```

**6** Now lets put the two sets together, using *rbind* to paste top-to-bottom:

```{r merging-sets}
data <- rbind(train_frame, test_frame)
```

**7** The column names actually contain many characters we do not want there. Let's clean:

```{r clean-up-names}
names       <- names(data)
names       <- gsub("[)(,\\-]","", names)
names(data) <- names 
```

**8** Now find the columns we want to keep:

Let us start by creating a list of all features with name containing *'mean'*, *'Mean'* or *'std'*:

```{r select-columns}
columns <- grep("[[Mm]ean|std]", features[,2]) 
columns <- columns +2       # accounting for the added activity and subject columns
columns <- c(1,2,columns)   # that subsequently need to be added!!
```

Now we are ready to strip our dataset, only keeping those values we are interested in:

```{r strip-data-set}
data <- data[,columns]
```

Dimensions should be 10299 by 53, check:
```{r check-size}
dd <- dim(data)
dd
```

So: 10299==`r dd[1]`   
And:   53==`r dd[2]`  


**9** Group!

```{r}
new_data_set <- group_by(data, activity, subject) %>% summarise_each(funs(mean))
```

And examine:

```{r}
str(new_data_set)
summary(new_data_set)
head(new_data_set)